---
title: "p8105_hw2_kb3066"
author: "Kaila Boyer"
date: "10/3/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1
Cleaning Mr.Trash Wheel sheet
```{r}
trash_wheel <- read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 1, range = "A2:N535") %>% janitor::clean_names()

trash_wheel <- trash_wheel %>% filter(!is.na(dumpster) & !is.na(year)) %>% 
                               mutate(sports_balls = round(sports_balls), 
                                      date = lubridate::ymd(date))

trash_wheel %>% filter(year == 2019) %>% summarise(med_sb = median(sports_balls))
```

Precipitation sheets 
```{r}
precip_2018 <- read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 7, 
                          range = "A2:B14") %>% janitor::clean_names()

precip_2019 <- read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 6, 
                          range = "A2:B14") %>% janitor::clean_names()
precip_2018 <- precip_2018 %>% mutate(year = "2018", 
                                      month = month.name)

precip_2019 <- precip_2019 %>% mutate(year = "2019", 
                                      month = month.name)
```
Merging precipitation data 
```{r}
precip_merged <- bind_rows(precip_2018, precip_2019)

precip_merged %>% group_by(year) %>% summarise(totals = sum(total)) %>% knitr::kable()
```

`In the Mr. Trashwheel dataset there are 453 observations and 14 columns. It includes variables on the various categories of litter and the amount collected. This includes data on cigarette butts, grocery bags, plastic bottles and sports balls. The number of homes powered through the collection of litter is another variable in this dataframe. The median number of sports balls collected by a dumpster in 2019 was 9.`
`The precipitation data contains variables for the monthly precipitation totals. The data frames from 2018 and 2019 had 12 observations, one for each month, along with a month and total precipitation variable. The total precipitation in 2018 was 70.33 inches. `


## Problem 2 

pols_month dataframe 
```{r}
pols_month <- read_csv("fivethirtyeight_datasets/pols-month.csv") %>% janitor::clean_names()
pols_month <- pols_month %>% 
                  separate(mon, into = c("year", "month", "day")) %>% 
                  mutate(month = rep(month.name, length.out = nrow(pols_month)), 
                         president = if_else(prez_gop == 1 & prez_dem !=1, "Republican", "Democrat"))

pols_month <- pols_month %>% select(!contains("prez"), -day) 
```

SNP dataframe
```{r}
snp <- read_csv("fivethirtyeight_datasets/snp.csv")
  
snp <-snp %>% separate(date, into = c("day", "month", "year")) %>% 
                  mutate(month = rep(month.name, length.out = nrow(snp)), 
                         year = as.numeric(year))  %>% 
                select(-day) 

snp <- snp %>% mutate(year = ifelse( year>=50 & year<=99, year+1900, 
                             ifelse(year>=00 & year<= 15, year+2000, NA)), 
                      year = as.character(year))
```

Unemployment dataframe
```{r}
unemployment <- read_csv("fivethirtyeight_datasets/unemployment.csv") %>% janitor::clean_names()
 
unemployment <- unemployment %>% pivot_longer(-year, names_to = "month", 
                                             values_to = "percent UE") %>% janitor::clean_names()


unemployment <- unemployment %>% mutate(month = rep(month.name, length.out = nrow(unemployment)), 
                                        year = as.character(year)) 
```

```{r}
pols_snp <- pols_month %>% left_join(snp, by = c("year", "month") )

fivethirtyeight_all <- pols_snp %>% left_join(unemployment, by = c("year", "month"))
```

`The pols_month datasets contains information on the number of elected officials in each political party for each day, starting in January 1947 until June 2015. This includes the number of Republicans and Democrats in office for President, Congress (Senators and Representatives), and Governors. This dataset has 822 observations and 9 variables.`
`The snp dataset has information on the closing values of the S&P index on a given date. It has two variables (date and closing value) and it has 787 observations.The time ranges from 1950 to 2015. `
`The unemployment dataset has information on the percent unemployment for each month. After transforming it to a longer format, it has 816 observations and 3 variables (year, month, and percent). The years range from 1948 to 2015.  `
`Merging all three datasets result in a larger dataframe with 822 observations and 11 variables. Like with pols_month, the earliest time provided is January 1947 and the latest is June 2015. The earliest date with closing S&P index values and a percent unemployment is 1950.`

## Problem 3 

Baby names data 
```{r}
baby_names <- read_csv("Popular_Baby_Names.csv") %>% 
                          janitor::clean_names() %>% 
                          distinct() %>% 
                          mutate(gender = str_to_lower(gender), 
                                 ethnicity = str_to_lower(ethnicity), 
                                 year_of_birth = as.character(year_of_birth), 
                                 childs_first_name = str_to_sentence(childs_first_name)) 

Hmisc::describe(baby_names$ethnicity)
baby_names <- baby_names %>% mutate(
                                   ethnicity = if_else(ethnicity == "asian and paci" | ethnicity == "asian and pacific islander", "Asian and Pacific Islander", 
                                                       if_else(ethnicity == "black non hisp" | ethnicity == "black non hispanic", "Black Non Hispanic", 
                                                               if_else(ethnicity == "white non hispanic" | ethnicity == "white non hisp", "White Non Hispanic", "Hispanic")))
)
```
Oliva dataframe 
```{r}
olivia <- baby_names %>% filter(childs_first_name == "Olivia") %>% 
  pivot_wider(names_from = year_of_birth, 
              values_from = c(rank, count)) %>% 
  mutate(ethnicity = as.factor(ethnicity)) %>% 
  select(-gender, -starts_with("count"))
  
```

```{r}
ethan <- baby_names %>% filter(gender == "male" & childs_first_name == "Ethan") %>% 
  pivot_wider(names_from = year_of_birth, 
              values_from = c(rank, count)) %>% 
  mutate(ethnicity = as.factor(ethnicity)) %>% 
  select(-gender, -starts_with("count"))
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
baby_names %>% filter(gender == "male" & ethnicity == "White Non Hispanic" & year_of_birth == 2016) %>% 
  ggplot(aes(x = rank, y = count)) + 
  geom_point() + 
  labs( 
    title = "Rank and Count of White Non Hispanic Male Names in 2016"
      )
```



